---
title: "LPI Transects + Quadrats Fall 2025 Data Clean-Up"
author: "Alexi C. Besser"
format:
  html:
    embed-resources: true
    code-background: true
    code-line-numbers: true
toc: true
editor: visual
---

# README

This code reads in individual csv files containing modified line point-intercept (LPI) + quadrat transect data collected at the Santa Rita Experimental Range (SRER), Onaqui (ONAQ), Moab (MOAB), and Jornada Experimental Range (JORN) National Ecological Observatory Network (NEON) sites in Spring 2025. GPS coordinates were taken using a SparkFun real-time kinematic (RTK) Express global navigation satellite system (GNSS) mounted to a 1.74 m monopole. Data for each transect were recorded in a unique project in the SW Maps Application on an iPad. Drop-down menus were used for primary fields and free-form text was used for "other" fields. Replicate data points and data points containing irreconcilable typos or other mistakes are removed in this code. Detailed descriptions of data collection methods and variables are included for each site.

A table of common plant codes can be found in the shared Google Drive and in ASU's LabArchives Notebook.

## Load Packages

```{r setup}

library(tidyr)
library(here)
library(readr)
library(plyr)
library(dplyr)
library(stringr)

```

# SRER

## 20-24 April 2025

**Personnel:** Alexi Besser, Isabel Torres, Heather Throop

**Brief Description of Modified LPI Protocol:** Transects ran S to N and were read on the W side. A pin flag was dropped every 1 m along the transect. All vegetation layers (including canopies) intersecting this pin were noted as dead or living and were recorded to species if known and life-form if not. The top layer height (cm), soil surface texture, litter type(s), litter depth (mm), and microsite were also recorded. A SparkFun RTK Express GNSS was used to record the location of each transect point.

**Brief Description of Litter Quadrat Protocol:** A 10 x 10 cm quadrat guide was placed on the ground with the pin flag in the SE corner every 5 m along the transect (except for transect SRER041; see below). A photograph of the quadrat was taken using the "narrow" setting on a GoPro Hero 9 Black outfitted with the standard lens followed by a photograph of the canopy taken using the "wide" setting on GoPro Hero 9 Black outfitted with a fisheye lens. Litter depths (mm) were measured in the center and each corner of the quadrat. All surface litter was collected from the quadrat using gloved hands, forceps, and paint brushes + mini dustpan as needed. In the lab, quadrat litter was dried, sorted, and weighed (mg).

50 m transects: SRER018, SRER021, SRER026, SRER034, SRER046

50 m transect with quadrats at 1 m, 2 m, 3 m, 4 m, 5 m, 34 m, 37 m, 46 m, 47 m, 48 m, 49 m, 50 m: SRER041

50 m transects with quadrats every 5 m (50 m, 45 m, 40 m, 35 m, 30 m, 25 m, 20 m, 15 m, 10 m, 5 m): SRER001, SRER030, SRER050, SRER039

### Levels 0 and 1 Data

```{r import_SRER_L0}

### LEVEL 0 DATA ###

# specify the Dropbox folder containing L0 transect data (individual csv files) and download it locally
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
# FOR NOW: copying and pasting this link in a web browser will automatically download all the csv files
SRER_L0_Dropbox_folder <- "https://www.dropbox.com/scl/fo/smaw1p81a1zxoknzrq7o9/AO0ry1SISFlULfHg8eVZ_mI?rlkey=xnljoexwgvjepso5hwsaqhnw5&st=ob1brlj2&dl=1"

# specify the local folder containing the downloaded L0 transect data
setwd("/Users/AlexiBesser/Downloads/SRER_LPI_L0_Spring_2025_csv_files")
SRER_L0_data_files <- dir("/Users/AlexiBesser/Downloads/SRER_LPI_L0_Spring_2025_csv_files")


### LEVEL 1 DATA ###

# read in L0 csv files and create a new column for transect ID
SRER_L1_data <- read_csv(SRER_L0_data_files, id = "Transect") %>%
   mutate(Transect = str_remove_all(Transect, ".csv"))

# write a csv file of the L1 transect data (compiled), save it locally, then upload it to the appropriate Dropbox folder
write.csv(SRER_L1_data, "/Users/AlexiBesser/Desktop/SRER_LPI_L1_Spring_2025_Data.csv")

```

There is now 1 csv file containing data for all 10 transects. This is considered level 1 data.

**L1 Dropbox Filepath:** ASU NSF Dryland Decomposition \> Field&LabCircle \> Field_Data \> SRER \> Transect_LPI_Data \> Level_1 \> Spring_2025 \> csv_file

```{r read_SRER_L1}

# read in L1 transect data
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
#SRER_L1_data <- read.csv("")

```

### Clean Level 1 Data

Investigate and rectify duplicate rows.

```{r SRER_merge_duplicates}

# create a data frame for duplicate points 
SRER_duplicate_points <- SRER_L1_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup() 

# there are 6 rows of duplicate points

# SRER018 (35): 2 rows with "35" for Remarks column ("17" and "18" in ID column). All the attributes for one of these rows (ID "18") are empty - this one will be removed from the data frame.
SRER_trimmed_data <- SRER_L1_data[!(SRER_L1_data$Transect=="SRER018" & SRER_L1_data$ID == "18"),]

# SRER034 (37): 2 rows with "37" for Remarks column ("14" and "16" in ID column). ID "16" contains the file name for a photograph ("20250422232557.jpg") - this file name will be copied and pasted into the Photo column for the ID "14" row and the ID "16" row will be deleted.
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER034" & SRER_trimmed_data$ID == "14", "Photo"] <- "20250422232557.jpg"
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER034" & SRER_trimmed_data$ID == "16"),]

# SRER039 (38): 2 rows with "38" for Remarks column ("14" and "23" in ID column). All the attributes are similar, so we will keep the later row, as there was an issue with the GPS around this point in the transect.
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "14"),]

# double-check that all duplicate points were fixed
SRER_duplicate_points_check <- SRER_trimmed_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup() 

```

Look at unique *Remarks*. Remove rows that are not points on transects but rather notes at the end of transects and rows with "NA" in every column.

```{r SRER_clean_Remarks}

# extract and review unique Remarks
unique(SRER_trimmed_data$Remarks)

# looks like we need to merge GPS coordinates with LPI data for SRER039 (there was an issue with the GPS)

# SRER039 (39): copy and paste GPS coordinates from the row for ID "22" into the row for ID "13"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "13", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "22", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "22"),]
# SRER039 (40): copy and paste GPS coordinates from the row for ID "21" into the row for ID "12"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "12", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "21", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "21"),]
# SRER039 (41): copy and paste GPS coordinates from the row for ID "20" into the row for ID "11"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "11", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "20", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "20"),]
# SRER039 (42): copy and paste GPS coordinates from the row for ID "19" into the row for ID "10"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "10", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "19", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "19"),]
# SRER039 (43): copy and paste GPS coordinates from the row for ID "18" into the row for ID "9"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "9", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "18", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "18"),] 
# SRER039 (44): copy and paste GPS coordinates from the row for ID "17" into the row for ID "16"
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "16", 4:18] <- SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "17", 4:18]
SRER_trimmed_data <- SRER_trimmed_data[!(SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "17"),]

# double-check that the Remarks column now only contains m points
unique(SRER_trimmed_data$Remarks)

```

There are now 500 rows in the trimmed data frame! 10 transects x 50 points per transect = 500 data points, so this makes sense, but double-check that there are 50 rows for every transect.

```{r SRER_check_rows_per_transect}

# create a character string of the transects
SRER_Transects <- unique(SRER_trimmed_data$Transect)

# create an empty list for the number of rows per transect
SRER_rows_per_transect <- list()

# run a for loop to count the number of rows per transect
for (i in SRER_Transects) {
  SRER_rows_per_transect[[i]] <- sum(SRER_trimmed_data$Transect == i)
  cat("Transect:", i, "- Rows:", SRER_rows_per_transect[[i]], "\n")
}

```

Investigate "NA" for *Top_layer* and look at unique values for *Top_layer_other* to check for typos.

```{r SRER_clean_Top_layer}

# create a data frame for rows with an "NA" for Top_layer
SRER_NA_Top_layer <- subset(SRER_trimmed_data, is.na(Top_layer))

# there are 4 rows that have an "NA" for Top_layer
# unfortunately, 3 of these rows also have "NA" for every other column and have to be removed from the data frame
# SRER030 (40): ID "12" - includes a quadrat - litter is visible in the quadrat photograph (GOPRO446.jpg) but the only columns filled out for this point are "0" for Litter_depth_west, "L-ISTE_patch" for Nearby_canopy_1, and "1012" for BagID, so this point also needs to be removed from the data frame
SRER_trimmed_data <- SRER_trimmed_data[!is.na(SRER_trimmed_data$Top_layer),]

# look at unique values for Top_layer_other
unique(SRER_trimmed_data$Top_layer_other)

# fix typos

# "PRVEA" appears once for transect SRER046 - this must be "PRVE"
SRER_trimmed_data$Top_layer_other[SRER_trimmed_data$Top_layer_other=="PRVEA"] <- "PRVE"

# "L-LATR" appears once for transect SRER050 - this should just be "LATR"
SRER_trimmed_data$Top_layer_other[SRER_trimmed_data$Top_layer_other=="L-LATR"] <- "LATR"

# "PAFL21" appears once for transect SRER021 - this should just be "PAFL" (the top height is 210, so it was probably initially typed into Top_layer_other rather than Top_height)
SRER_trimmed_data$Top_layer_other[SRER_trimmed_data$Top_layer_other=="PAFL21"] <- "PAFL"

```

```{r SRER_clean_Lower_1}

# create a data frame for rows with an "NA" for Lower_1
SRER_NA_Lower_1 <- subset(SRER_trimmed_data, is.na(Lower_1))

# there are 5 rows that have an "NA" for Lower_1
# these rows all have values in the Top_layer and/or Top_layer_other column and have "N" values for Lower_2 and Lower_3, so we will assume these "NA" values should all be "N"
SRER_trimmed_data$Lower_1[is.na(SRER_trimmed_data$Lower_1)] <- "N"

# look at unique values for Lower_1_other
unique(SRER_trimmed_data$Lower_1_other)

# no typos

```

```{r SRER_clean_Lower_2}

# create a data frame for rows with an "NA" for Lower_2
SRER_NA_Lower_2 <- subset(SRER_trimmed_data, is.na(Lower_2))

# there are no rows with an "NA" for Lower_2

# look at unique values for Lower_2_other
unique(SRER_trimmed_data$Lower_2_other)

# no typos

```

```{r SRER_clean_Lower_3}

# create a data frame for rows with an "NA" for Lower_3
SRER_NA_Lower_3 <- subset(SRER_trimmed_data, is.na(Lower_3))

# there is 1 row with an "NA" for Lower_3
# SRER021 (40): ID "11" - all other layers have an "N", so we can safely assume Lower_3 should be "N", too
SRER_trimmed_data$Lower_3[is.na(SRER_trimmed_data$Lower_3)] <- "N"

# look at unique values for Lower_3_other
unique(SRER_trimmed_data$Lower_3_other)

# no typos

```

```{r SRER_clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
SRER_NA_Soil_surface <- subset(SRER_trimmed_data, is.na(Soil_surface))

# there are 2 rows that have an "NA" for Soil_surface

# the points on either side of both of these points have "SOIL" for Soil_surface, so we will go ahead and assume these should be "SOIL" as well
SRER_trimmed_data$Soil_surface[is.na(SRER_trimmed_data$Soil_surface)] <- "SOIL"

# look at unique values for Soil_surface_other
unique(SRER_trimmed_data$Soil_surface_other)

# no typos

```

```{r SRER_clean_Microsite}

# create a data frame for rows with an "NA" for Microsite
SRER_NA_Microsite <- subset(SRER_trimmed_data, is.na(Microsite))

# there are 3 rows that have an "NA" for Microsite

# 2 of these rows have "N" for Top_layer and 1 has "D-FORB" for Top_layer, so they should all have "OPEN" for Microsite
SRER_trimmed_data$Microsite[is.na(SRER_trimmed_data$Microsite)] <- "OPEN"

# look at unique values for Microstie_other
unique(SRER_trimmed_data$Microsite_other)

# no typos

```

```{r SRER_clean_Top_height}

# create a data frame for rows with an "NA" for Microsite
SRER_NA_Top_height <- subset(SRER_trimmed_data, is.na(Top_height))

# there are 6 rows that have an "NA" for Top_height

# 1 of these rows has an "L-OTHER" for Top_layer and is surrounded by points that have the same Top_layer and a Top_height of 60, so we will assume the Top_height should be 60
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER041" & SRER_trimmed_data$ID == "4", "Top_height"] <- 60

# the rest should be 0
SRER_trimmed_data$Top_height[is.na(SRER_trimmed_data$Top_height)] <- 0

```

```{r SRER_clean_Litter}

# create a data frame for rows with an "NA" for Litter
SRER_NA_Litter <- subset(SRER_trimmed_data, is.na(Litter))

# there are no rows that have an "NA" for Litter

# look at unique values for Litter_other
unique(SRER_trimmed_data$Litter_other)

# no typos
# create a data frame for rows with an "NA" for Litter
SRER_NA_Litter_depth <- subset(SRER_trimmed_data,
  is.na(Litter_depth_center) |
  is.na(Litter_depth_north) |
  is.na(Litter_depth_east) |
  is.na(Litter_depth_south) |
  is.na(Litter_depth_west))

# 1 of these rows has "N" for Litter
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER050" & SRER_trimmed_data$ID == "29", "Litter_depth_north"] <- "0"

# 1 of these rows has a quadrat photo and there looks to be a pretty even distribution of litter - let's take the average of the other 4 litter depth measurements
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER039" & SRER_trimmed_data$ID == "37", "Litter_depth_west"] <- 4

# let's just take the average of the other 4 litter depth measurements
SRER_trimmed_data[SRER_trimmed_data$Transect=="SRER050" & SRER_trimmed_data$ID == "15", "Litter_depth_west"] <- 2

```

```{r SRER_clean_Nearby_canopy}

# look at unique values for Nearby_canopy_1
unique(SRER_trimmed_data$Nearby_canopy_1)

# fix typos
SRER_trimmed_data$Nearby_canopy_1[SRER_trimmed_data$Nearby_canopy_1 == "ERLE_DSUBS_patch"] <- "ERLE_D-SUBS_patch"

# look at unique values for Nearby_canopy_2
unique(SRER_trimmed_data$Nearby_canopy_2)

# fix typos
SRER_trimmed_data$Nearby_canopy_2[SRER_trimmed_data$Nearby_canopy_2 == "L-Op_95SW_25H_35L_15P"] <- "L-OP_95SW_25H_35L_15P"

# look at unique values for Nearby_canopy_3
unique(SRER_trimmed_data$Nearby_canopy_3)

# look at unique values for Nearby_canopy_4
unique(SRER_trimmed_data$Nearby_canopy_4)

# look at unique values for Nearby_canopy_5
unique(SRER_trimmed_data$Nearby_canopy_5)

```

### Level 2 Data

```{r save_SRER_L2}

# write a csv file of the L2 transect data, save it locally, then upload it to the appropriate Dropbox folder
write.csv(SRER_trimmed_data, "/Users/AlexiBesser/Desktop/SRER_LPI_L2_Spring_2025_Data.csv")

```

### Transect Locations

```{r SRER_transect_locations}

SRER_Spring_2025_Transect_Locations <- SRER_trimmed_data %>%
  filter(Remarks == 25) %>%
  select(Transect, Latitude, Longitude, Elevation)

```

### Quadrat Sample Log

```{r quadrat_sample_log}

SRER_quad_samp_log <- SRER_trimmed_data %>%
  select(BagID, Transect, Remarks, Litter, Litter_other, Litter_depth_center,
         Litter_depth_north, Litter_depth_east, Litter_depth_south,
         Litter_depth_west, Nearby_canopy_1, Nearby_canopy_2, Nearby_canopy_3,
         Nearby_canopy_4, Nearby_canopy_5)

SRER_quad_samp_log <- SRER_quad_samp_log[!(is.na(SRER_quad_samp_log$BagID)),]

write.csv(SRER_quad_samp_log, "SRER_Litter_Quadrat_Sample_Log_Spring_2025.csv", row.names = FALSE)

```

# JORN

## 25-29 April 2025

**Personnel:** Alexi Besser, Isabel Torres

**Brief Description of Modified LPI Protocol:** Transects ran S to N and were read on the W side. A pin flag was dropped every 1 m along the transect. All vegetation layers (including canopies) intersecting this pin were noted as dead or living and were recorded to species if known and life-form if not. The top layer height (cm), soil surface texture, litter type(s), litter depth (mm), and microsite were also recorded. A SparkFun RTK Express GNSS was used to record the location of each transect point.

50 m transects: JORN102, JORN104, JORN105, JORN109, JORN111, JORN114, JORN117, JORN120, JORN126, JORN138, JORN139, JORN140, JORN141, JORN144

50 m transects with quadrats every 5 m (50 m, 45 m, 40 m, 35 m, 30 m, 25 m, 20 m, 15 m, 10 m, 5 m): JORN019, JORN035, JORN037, JORN123, JORN129, JORN137

### Levels 0 and 1 Data

```{r import_JORN_L0}

### LEVEL 0 DATA ###

# specify the Dropbox folder containing L0 transect data (individual csv files) and download it locally
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
# FOR NOW: copying and pasting this link in a web browser will automatically download all the csv files
JORN_L0_Dropbox_folder <- "https://www.dropbox.com/scl/fo/lzbc5lp259cl4b9uukf9n/AGPCTbNu63Ifj74PO3JMxeE?rlkey=30r5xyy71xqbmby838qob5vh5&st=vmotyt9a&dl=1"

# specify the local folder containing the downloaded L0 transect data
setwd("/Users/AlexiBesser/Downloads/JORN_LPI_L0_Spring_2025_csv_files")
JORN_L0_data_files <- dir("/Users/AlexiBesser/Downloads/JORN_LPI_L0_Spring_2025_csv_files")


### LEVEL 1 DATA ###

# read in L0 csv files and create a new column for transect ID
JORN_L1_data <- read_csv(JORN_L0_data_files, id = "Transect") %>%
   mutate(Transect = str_remove_all(Transect, ".csv"))

# write a csv file of the L1 transect data (compiled), save it locally, then upload it to the appropriate Dropbox folder
write.csv(JORN_L1_data, "/Users/AlexiBesser/Desktop/JORN_LPI_L1_Spring_2025_Data.csv")

```

There is now 1 csv file containing data for all 20 transects. This is considered level 1 data.

**L1 Dropbox Filepath:** ASU NSF Dryland Decomposition \> Field&LabCircle \> Field_Data \> JORN \> Transect_LPI_Data \> Level_1 \> Spring_2025 \> csv_file

```{r read_JORN_L1}

# read in L1 transect data
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
#SRER_L1_data <- read.csv("")

```

### Clean Level 1 Data

Investigate and rectify duplicate rows.

```{r JORN_merge_duplicates}

# create a data frame for duplicate points 
JORN_duplicate_points <- JORN_L1_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup() 

# there are 12 rows of duplicate points

# JORN037 (17): 2 rows with "17" for Remarks column ("35" and "36" in ID column). All the attributes for one of these rows (ID "36") are empty - this one will be removed from the data frame.
JORN_trimmed_data <- JORN_L1_data[!(JORN_L1_data$Transect=="JORN037" & JORN_L1_data$ID == "36"),]

# JORN105 (36): 2 rows with "36" for Remarks column ("15" and "16" in ID column). All the attributes for one of these rows (ID "16") are empty except for Litter_depth_west, which is "0" - this value will be pasted into the Litter_depth_west column for the ID "15" row and the ID "16" row will be deleted.
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN105" & JORN_trimmed_data$ID == "15", "Litter_depth_west"] <- "0"
JORN_trimmed_data <- JORN_trimmed_data[!(JORN_trimmed_data$Transect=="JORN105" & JORN_trimmed_data$ID == "16"),]

# JORN105 (29): 2 rows with "29" for Remarks column ("23" and "24" in ID column). All the attributes for one of these rows (ID "14") are empty except for Litter_depth_west, which is "0" - this value will be pasted into the Litter_depth_west column for the ID "23" row and the ID "24" row will be deleted.
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN105" & JORN_trimmed_data$ID == "23", "Litter_depth_west"] <- "0"
JORN_trimmed_data <- JORN_trimmed_data[!(JORN_trimmed_data$Transect=="JORN105" & JORN_trimmed_data$ID == "24"),]

# JORN117 (14): 2 rows with "14" for Remarks column ("37" and "38" in ID column). All the attributes for one of these rows (ID "38") are empty - this one will be removed from the data frame.
JORN_trimmed_data <- JORN_trimmed_data[!(JORN_trimmed_data$Transect=="JORN117" & JORN_trimmed_data$ID == "38"),]

# JORN138 (25): 2 rows with "25" for Remarks column ("26" and "27" in ID column). All the attributes for one of these rows (ID "27") are empty - this one will be removed from the data frame.
JORN_trimmed_data <- JORN_trimmed_data[!(JORN_trimmed_data$Transect=="JORN138" & JORN_trimmed_data$ID == "27"),]

# JORN141 (37): 2 rows with "37" for Remarks column ("14" and "16" in ID column). The row with the ID "16" should be for the 35 m mark, which is missing.
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN141" & JORN_trimmed_data$ID == "16", "Remarks"] <- "35"

# double-check that all duplicate points were fixed
JORN_duplicate_points_check <- JORN_trimmed_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup() 

```

Look at unique *Remarks*. Remove rows that are not points on transects but rather notes at the end of transects and rows with "NA" in every column.

```{r JORN_clean_Remarks}

# extract and review unique Remarks
unique(JORN_trimmed_data$Remarks)

# pull dune start and end GPS coordinates into a separate data frame
dune_GPS_coords <- JORN_trimmed_data[grepl("[A-Za-z]", JORN_trimmed_data$Remarks), ]

# write a csv file for the dune start and end GPS coordinates
write.csv(dune_GPS_coords, "JORN_dune_GPS_coords_Spring_2025.csv", row.names = FALSE)

# make a character string of the Remarks for dune start and end GPS coordinate rows
dune_GPS_remarks <- dune_GPS_coords$Remarks

# remove dune start and end GPS coordinates from the main data frame
JORN_trimmed_data <- JORN_trimmed_data[!grepl(paste(dune_GPS_remarks, collapse = "|"), JORN_trimmed_data$Remarks), ]

# double-check that the Remarks column now only contains m points
unique(JORN_trimmed_data$Remarks)

```

There are now 1,000 rows in the trimmed data frame! 20 transects x 50 points per transect = 1,000 data points, so this makes sense, but double-check that there are 50 rows for every transect.

```{r JORN_check_rows_per_transect}

# create a character string of the transects
JORN_Transects <- unique(JORN_trimmed_data$Transect)

# create an empty list for the number of rows per transect
JORN_rows_per_transect <- list()

# run a for loop to count the number of rows per transect
for (i in JORN_Transects) {
  JORN_rows_per_transect[[i]] <- sum(JORN_trimmed_data$Transect == i)
  cat("Transect:", i, "- Rows:", JORN_rows_per_transect[[i]], "\n")
}

```

Investigate "NA" for *Top_layer* and look at unique values for *Top_layer_other* to check for typos.

```{r JORN_clean_Top_layer}

# create a data frame for rows with an "NA" for Top_layer
JORN_NA_Top_layer <- subset(JORN_trimmed_data, is.na(Top_layer))

# there are 15 rows that have an "NA" for Top_layer

# 7 of these rows have an "N" for Lower_layer_1 and a Top_height of 0, so we will assume their Top_layer should be "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN104" & JORN_trimmed_data$ID == "12", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN105" & JORN_trimmed_data$ID == "52", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN111" & JORN_trimmed_data$ID == "13", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN111" & JORN_trimmed_data$ID == "22", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN111" & JORN_trimmed_data$ID == "23", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN123" & JORN_trimmed_data$ID == "42", "Top_layer"] <- "N"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN139" & JORN_trimmed_data$ID == "6", "Top_layer"] <- "N"

# JORN037 (18): ID "34" has a Lower_layer_1 of "D-PGRASS" and a Top_height of 10, but not other lower layers - let's assume the Top_layer should be D-PGRASS and there are no lower layers
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN037" & JORN_trimmed_data$ID == "34", "Top_layer"] <- "D-PGRASS"
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN037" & JORN_trimmed_data$ID == "34", "Lower_1"] <- "N"

# unfortunately, the 7 remaining rows also have "NA" for every other column (except Litter_depth_west, which is 0 for 6/7 rows) and have to be removed from the data frame
# first, put the 5 rows that had quadrats in a separate data frame to include in the quadrat sample log later one
JORN_NA_quadrats <- JORN_trimmed_data[is.na(JORN_trimmed_data$Top_layer),]
JORN_NA_quadrats <- JORN_NA_quadrats %>%
  filter(ID %in% c(35, 46, 38, 50, 22))
# now, remove the rows from the data frame
JORN_trimmed_data <- JORN_trimmed_data[!is.na(JORN_trimmed_data$Top_layer),]

# look at unique values for Top_layer_other
unique(JORN_trimmed_data$Top_layer_other)

# no typos


```

```{r JORN_clean_Lower_1}

# create a data frame for rows with an "NA" for Lower_1
JORN_NA_Lower_1 <- subset(JORN_trimmed_data, is.na(Lower_1))

# there are 11 rows that have an "NA" for Lower_1
# these rows all have values in the Top_layer and/or Top_layer_other column and have "N" values for Lower_2 and Lower_3, so we will assume these "NA" values should all be "N"
JORN_trimmed_data$Lower_1[is.na(JORN_trimmed_data$Lower_1)] <- "N"

# look at unique values for Lower_1_other
unique(JORN_trimmed_data$Lower_1_other)

# fix typos

# "PRGO" appears once for transect JORN019 - this should just be "PRGL"
JORN_trimmed_data$Lower_1_other[JORN_trimmed_data$Lower_1_other=="PRGO"] <- "PRGL"

```

```{r JORN_clean_Lower_2}

# create a data frame for rows with an "NA" for Lower_2
JORN_NA_Lower_2 <- subset(JORN_trimmed_data, is.na(Lower_2))

# there are 6 rows with an "NA" for Lower_2

# all of these rows have an "N" for Lower_1, so Lower_2 should be "N" as well
JORN_trimmed_data$Lower_2[is.na(JORN_trimmed_data$Lower_2)] <- "N"

# look at unique values for Lower_2_other
unique(JORN_trimmed_data$Lower_2_other)

# no typos

```

```{r JORN_clean_Lower_3}

# create a data frame for rows with an "NA" for Lower_3
JORN_NA_Lower_3 <- subset(JORN_trimmed_data, is.na(Lower_3))

# there are 9 rows with an "NA" for Lower_3
# all of these rows have an "N" for Lower_2, so Lower_3 should be "N" as well
JORN_trimmed_data$Lower_3[is.na(JORN_trimmed_data$Lower_3)] <- "N"

# look at unique values for Lower_3_other
unique(JORN_trimmed_data$Lower_3_other)

# no typos

```

```{r JORN_clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
JORN_NA_Soil_surface <- subset(JORN_trimmed_data, is.na(Soil_surface))

# there are 3 rows that have an "NA" for Soil_surface

# all points along the transect for one of these points have "SOIL" for Soil_surface, so we will go ahead and assume this point should have "SOIL" as well
# the other two points are from the same transect, the point in the middle of them has "BURROW" for Soil_surface, but the vast majority of the other points have "SOIL", so we will go ahead and assume this point should have "SOIL" as well
JORN_trimmed_data$Soil_surface[is.na(JORN_trimmed_data$Soil_surface)] <- "SOIL"

# look at unique values for Soil_surface_other
unique(JORN_trimmed_data$Soil_surface_other)

# no typos

```

d

```{r JORN_clean_Microsite}

# create a data frame for rows with an "NA" for Microsite
JORN_NA_Microsite <- subset(JORN_trimmed_data, is.na(Microsite))

# there are 9 rows that have an "NA" for Microsite

# all of these rows have "N" for Top_layer and a Top_height of 0, so they should all have "OPEN" for Microsite
JORN_trimmed_data$Microsite[is.na(JORN_trimmed_data$Microsite)] <- "OPEN"

# look at unique values for Microstie_other
unique(JORN_trimmed_data$Microsite_other)

# no typos

```

```{r JORN_clean_Top_height}

# create a data frame for rows with an "NA" for Microsite
JORN_NA_Top_height <- subset(JORN_trimmed_data, is.na(Top_height))

# there are 5 rows that have an "NA" for Top_height

# all of these rows have "N" for Top_layer, so their Top_height should be 0
JORN_trimmed_data$Top_height[is.na(JORN_trimmed_data$Top_height)] <- 0

```

```{r JORN_clean_Litter}

# create a data frame for rows with an "NA" for Litter
JORN_NA_Litter <- subset(JORN_trimmed_data, is.na(Litter))

# there are 3 rows that have an "NA" for Litter

# they all have litter depth measurements of 0, so they should have "N" for Litter
JORN_trimmed_data$Litter[is.na(JORN_trimmed_data$Litter)] <- "N"

# look at unique values for Litter_other
unique(JORN_trimmed_data$Litter_other)

# no typos

# create a data frame for rows with "NA" values for litter depth measurements
JORN_NA_Litter_depth <- subset(JORN_trimmed_data,
  is.na(Litter_depth_center) |
  is.na(Litter_depth_north) |
  is.na(Litter_depth_east) |
  is.na(Litter_depth_south) |
  is.na(Litter_depth_west))

# there are 25 rows that have "NA" for at least one litter depth measurement

# make sure litter depth values are numeric
JORN_trimmed_data$Litter_depth_center <- as.numeric(JORN_trimmed_data$Litter_depth_center)
JORN_trimmed_data$Litter_depth_north <- as.numeric(JORN_trimmed_data$Litter_depth_north)
JORN_trimmed_data$Litter_depth_east <- as.numeric(JORN_trimmed_data$Litter_depth_east)
JORN_trimmed_data$Litter_depth_south <- as.numeric(JORN_trimmed_data$Litter_depth_south)
JORN_trimmed_data$Litter_depth_west <- as.numeric(JORN_trimmed_data$Litter_depth_west)

# 6 of these rows have "LT-H" for Litter or "H_W" for Litter_other, so we will take the average litter depth from the other four measurements and paste it into the missing measurement
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN037" & JORN_trimmed_data$ID == "42", "Litter_depth_west"] <- 0
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN117" & JORN_trimmed_data$ID == "26", "Litter_depth_west"] <- 0
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN126" & JORN_trimmed_data$ID == "42", "Litter_depth_north"] <- 1
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN137" & JORN_trimmed_data$ID == "2", "Litter_depth_north"] <- 1
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN111" & JORN_trimmed_data$ID == "32", "Litter_depth_center"] <- 1
JORN_trimmed_data[JORN_trimmed_data$Transect=="JORN129" & JORN_trimmed_data$ID == "3", "Litter_depth_north"] <- 2

# 1 row has "LT-W" for Litter and 1 has "B" for Litter_other, so these litter depth measurements should be 0
# the remaining 17 rows have "N" for Litter
JORN_trimmed_data$Litter_depth_center[is.na(JORN_trimmed_data$Litter_depth_center)] <- "0"
JORN_trimmed_data$Litter_depth_north[is.na(JORN_trimmed_data$Litter_depth_north)] <- "0"
JORN_trimmed_data$Litter_depth_east[is.na(JORN_trimmed_data$Litter_depth_east)] <- "0"
JORN_trimmed_data$Litter_depth_south[is.na(JORN_trimmed_data$Litter_depth_south)] <- "0"
JORN_trimmed_data$Litter_depth_west[is.na(JORN_trimmed_data$Litter_depth_west)] <- "0"

```

```{r JORN_clean_Nearby_canopy}

# look at unique values for Nearby_canopy_1
unique(JORN_trimmed_data$Nearby_canopy_1)

# fix typos
JORN_trimmed_data$Nearby_canopy_1[JORN_trimmed_data$Nearby_canopy_1 == "L-PRGL_DUNE"] <- "L-PRGL_dune"
JORN_trimmed_data$Nearby_canopy_1[JORN_trimmed_data$Nearby_canopy_1 == "L-PRGLE_80SE_20H_40L_30P"] <- "L-PRGL_80SE_20H_40L_30P"
JORN_trimmed_data$Nearby_canopy_1[JORN_trimmed_data$Nearby_canopy_1 == "L-PRGLE_patch_NW"] <- "L-PRGL_patch_NW"

# look at unique values for Nearby_canopy_2
unique(JORN_trimmed_data$Nearby_canopy_2)

# fix typos
JORN_trimmed_data$Nearby_canopy_2[JORN_trimmed_data$Nearby_canopy_2 == "L-PRGLE_25SE_45H_110L_100P"] <- "L-PRGL_25SE_45H_110L_100P"

# look at unique values for Nearby_canopy_3
unique(JORN_trimmed_data$Nearby_canopy_3)

# look at unique values for Nearby_canopy_4
unique(JORN_trimmed_data$Nearby_canopy_4)

# look at unique values for Nearby_canopy_5
unique(JORN_trimmed_data$Nearby_canopy_5)

```

### Level 2 Data

```{r save_JORN_L2}

# write a csv file of the L2 transect data, save it locally, then upload it to the appropriate Dropbox folder
write.csv(JORN_trimmed_data, "/Users/AlexiBesser/Desktop/JORN_LPI_L2_Spring_2025_Data.csv")

```

### Transect Locations

```{r JORN_transect_locations}

JORN_Spring_2025_Transect_Locations <- JORN_trimmed_data %>%
  filter(Remarks == 25) %>%
  select(Transect, Latitude, Longitude, Elevation)

```

### Quadrat Sample Log

```{r quadrat_sample_log}

JORN_quad_samp_log <- JORN_trimmed_data %>%
  select(BagID, Transect, Remarks, Litter, Litter_other, Litter_depth_center,
         Litter_depth_north, Litter_depth_east, Litter_depth_south,
         Litter_depth_west, Nearby_canopy_1, Nearby_canopy_2, Nearby_canopy_3,
         Nearby_canopy_4, Nearby_canopy_5)

JORN_quad_samp_log <- JORN_quad_samp_log[!(is.na(JORN_quad_samp_log$BagID)),]

JORN_NA_quad_samp_log <- JORN_NA_quadrats %>%
  select(BagID, Transect, Remarks, Litter, Litter_other, Litter_depth_center,
         Litter_depth_north, Litter_depth_east, Litter_depth_south,
         Litter_depth_west, Nearby_canopy_1, Nearby_canopy_2, Nearby_canopy_3,
         Nearby_canopy_4, Nearby_canopy_5)

JORN_quad_samp_log <- rbind(JORN_quad_samp_log, JORN_NA_quad_samp_log)

write.csv(JORN_quad_samp_log, "JORN_Litter_Quadrat_Sample_Log_Spring_2025.csv", row.names = FALSE)

```

# MOAB

## 15-19 May 2025

**Personnel:** Alexi Besser, Isabel Torres, Dellena Bloom, Allelua Niyokwizera, Heather Throop, Alex Cueva

**Brief Description of Modified LPI Protocol:** Transects ran S to N and were read on the W side. A pin flag was dropped every 1 m along the transect. All vegetation layers (including canopies) intersecting this pin were noted as dead or living and were recorded to species if known and life-form if not. The top layer height (cm), soil surface texture, litter type(s), litter depth (mm), and microsite were also recorded. A SparkFun RTK Express GNSS was used to record the location of each transect point.

50 m transects: MOAB104, MOAB106, MOAB114, MOAB116, MOAB118, MOAB119, MOAB122, MOAB126, MOAB127, MOAB135, MOAB147, MOAB148, MOAB149

50 m transects with quadrats every 5 m (50 m, 45 m, 40 m, 35 m, 30 m, 25 m, 20 m, 15 m, 10 m, 5 m): MOAB103, MOAB105, MOAB107, MOAB111, MOAB128, MOAB131, MOAB137, MOAB150

### Levels 0 and 1 Data

```{r MOAB}

### LEVEL 0 DATA ###

# specify the Dropbox folder containing L0 transect data (individual csv files) and download it locally
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
# FOR NOW: copying and pasting this link in a web browser will automatically download all the csv files
MOAB_L0_Dropbox_folder <- "https://www.dropbox.com/scl/fo/r11axu2km65ein82yr71i/ANi_MnnRHHhEYmnwgjDYEQw?rlkey=t043n6k9zhqew2xgk3kig0ocr&st=riw2t9qn&dl=1"

# Alexi: specify the local folder containing the downloaded L0 transect data
setwd("/Users/AlexiBesser/Downloads/MOAB_LPI_L0_Spring_2025_csv_files")
MOAB_L0_data_files <- dir("/Users/AlexiBesser/Downloads/MOAB_LPI_L0_Spring_2025_csv_files")

# Dellena
#MOAB_L0_data_files <- dir("Spring_2025/csv_files/")

# Dellena: read in L0 csv files and create a new column for transect ID
#MOAB_L1_data <- read_csv(file.path("Spring_2025/csv_files/",
#                                   MOAB_L0_data_files), id = "Transect") %>%
#  mutate(Transect = str_remove_all(Transect, ".csv")) %>%
#  mutate(Transect = str_remove_all(Transect, "_1")) %>%
#  mutate(Transect = str_remove_all(Transect, "_2"))

### LEVEL 1 DATA ###

# read in L0 csv files and create a new column for transect ID
MOAB_L1_data <- read_csv(MOAB_L0_data_files, id = "Transect") %>%
  mutate(Transect = str_remove_all(Transect, ".csv")) %>%
  mutate(Transect = str_remove_all(Transect, "_1")) %>%
  mutate(Transect = str_remove_all(Transect, "_2"))


```

There is now 1 csv file containing data for all 21 transects. This is considered level 1 data.

**L1 Dropbox Filepath:** ASU NSF Dryland Decomposition \> Field&LabCircle \> Field_Data \> MOAB \> Transect_LPI_Data \> Level_1 \> Spring_2025 \> csv_file

```{r read_MOAB_L1}

# read in L1 transect data
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
#SRER_L1_data <- read.csv("")

```

### Clean Level 1 Data

Investigate and rectify duplicate rows.

```{r MOAB_merge_duplicates}

# create a data frame for duplicate points 
MOAB_duplicate_points <- MOAB_L1_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup()

# there are 121 rows of duplicate points
# MOAB103 (45): 2 rows with "45" for Remarks column ("6" and "37" in ID column). One of these was a typo and should be "15"
MOAB_trimmed_data <- MOAB_L1_data
  
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB103" & MOAB_trimmed_data$ID == "37", "Remarks"] <- 15

MOAB_duplicate_points <- MOAB_duplicate_points[!(MOAB_duplicate_points$Transect=="MOAB103"),]

# MOAB135 (38): 2 rows with "38" for Remarks column ("13" and "15" in ID column). One of these was a typo and should be "36"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$ID == "15" & MOAB_trimmed_data$Remarks == "38", "Remarks"] <- 36

MOAB_duplicate_points <- MOAB_duplicate_points[!(MOAB_duplicate_points$Transect=="MOAB135"),]


# Creating separate data-frames for satellite data, where satellite height is 1.74 m and the plant information data-frame where satellite height is 0 m and then binding them together, removing the rows of NA - written by Dellena

# NOTE: T 103 remark 45 of the heights == 1.74 had writing that gets removed with first negative selection below, this also creates 2 ID columns, from both satellites and time columns, chose which you want or if you want to name them different things - denoted by .x and .y

MOAB_dup_NAcols_removed <- MOAB_duplicate_points %>%
  filter(`Instrument Ht` == 1.74) %>%
  select(-c(Top_layer:Quadrat_checklist)) %>%
  merge(MOAB_duplicate_points %>%
    filter(`Instrument Ht` == 0) %>%
    select(-c(Geometry:`Ortho Height`, `Fix ID`:`Satellites in Use`)), 
    by = c("Transect", "Remarks"))

colnames(MOAB_dup_NAcols_removed)[colnames(MOAB_dup_NAcols_removed) == "ID.x"] <- "ID"

colnames(MOAB_dup_NAcols_removed)[colnames(MOAB_dup_NAcols_removed) == "Time.x"] <- "Time"

colnames(MOAB_dup_NAcols_removed)[colnames(MOAB_dup_NAcols_removed) == "Instrument Ht.x"] <- "Instrument Ht"

MOAB_dup_NAcols_removed <- MOAB_dup_NAcols_removed %>%
    select(-c("ID.y", "Time.y", "Instrument Ht.y"))

MOAB_trimmed_data <- MOAB_trimmed_data %>%
  filter(!(`Instrument Ht` == 0))

MOAB_trimmed_data <- MOAB_trimmed_data %>%
  anti_join(MOAB_dup_NAcols_removed, by = c("Transect", "Remarks")) %>%
  bind_rows(MOAB_dup_NAcols_removed)

# create a data frame for duplicate points 
MOAB_duplicate_points_check <- MOAB_trimmed_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup()

# MOAB119 (8): 2 rows with "8" for Remarks column ("46" and "47" in ID column). They are identical. Remove the first one.
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "46"),]

# MOAB119 (11): 2 rows with "11" for Remarks column ("42" and "43" in ID column). They are identical. Remove the first one.
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "42"),]

```

Look at unique *Remarks*. Remove rows that are not points on transects but rather notes at the end of transects and rows with "NA" in every column.

```{r MOAB_clean_Remarks}

# extract and review unique Remarks
unique(MOAB_trimmed_data$Remarks)

# there is a "111" and a "118"

# "118" should be "18"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB118" & MOAB_trimmed_data$ID == "18", "Remarks"] <- 18

# I think "111" should be "1"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB111" & MOAB_trimmed_data$ID == "13", "Remarks"] <- 1

```

There are now 1,050 rows in the trimmed data frame! 21 transects x 50 points per transect = 1,050 data points, so this makes sense, but double-check that there are 50 rows for every transect.

```{r MOAB_check_rows_per_transect}

# create a character string of the transects
MOAB_Transects <- unique(MOAB_trimmed_data$Transect)

# create an empty list for the number of rows per transect
MOAB_rows_per_transect <- list()

# run a for loop to count the number of rows per transect
for (i in MOAB_Transects) {
  MOAB_rows_per_transect[[i]] <- sum(MOAB_trimmed_data$Transect == i)
  cat("Transect:", i, "- Rows:", MOAB_rows_per_transect[[i]], "\n")
}

# we are missing data for the 14 m mark on transect MOAB116 and 27 m mark on transect MOAB135

```

Investigate "NA" for *Top_layer* and look at unique values for *Top_layer_other* to check for typos.

```{r MOAB_clean_Top_layer}

# create a data frame for rows with an "NA" for Top_layer
MOAB_NA_Top_layer <- subset(MOAB_trimmed_data, is.na(Top_layer))

# there are 30 rows that have an "NA" for Top_layer

# all of these rows have "N" for lower layers, "OPEN" for Microsite, and 0 for Top_height, so we will assume their Top_layer should be "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB111" & MOAB_trimmed_data$ID == "33", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB114" & MOAB_trimmed_data$ID == "22", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB116" & MOAB_trimmed_data$ID == "17", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "44", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB149" & MOAB_trimmed_data$ID == "5", "Top_layer"] <- "N"

# all of these rows have 0 for Top_height, so we will assume their Top_layer should be "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "52", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "50", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "47", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "45", "Top_layer"] <- "N"

# expeditated data collection on transect MOAB135 due to storm, all of these Top_layer columns should be "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "14", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "15", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "16", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "17", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "18", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "19", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "20", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "22", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "30", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "32", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "33", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "34", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "35", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "36", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "37", "Top_layer"] <- "N"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "45", "Top_layer"] <- "N"

# all columns are empty for these rows, so they need to be deleted from the data frame
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$ID == "43"),]
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB126" & MOAB_trimmed_data$ID == "18"),]
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB128" & MOAB_trimmed_data$ID == "33"),]
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB131" & MOAB_trimmed_data$ID == "2"),]
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB147" & MOAB_trimmed_data$ID == "33"),]

# look at unique values for Top_layer_other
unique(MOAB_trimmed_data$Top_layer_other)

# no typos

```

```{r MOAB_clean_Lower_1}

# create a data frame for rows with an "NA" for Lower_1
MOAB_NA_Lower_1 <- subset(MOAB_trimmed_data, is.na(Lower_1))

# there are 34 rows that have an "NA" for Lower_1
# 29 of these rows have "N" for Top_layer, 2 have "L-PGRASS" for Top_layer, 2 have "L-OTHER" for Top_layer and "BOGR" for Top_layer_other, and 1 has "L-OTHER" for Top_layer and "ACHY" for Top_layer_other 
# several rows have "N" for random other lower layers
# we are going to assume all of these should be "N" for Lower_1
MOAB_trimmed_data$Lower_1[is.na(MOAB_trimmed_data$Lower_1)] <- "N"

# look at unique values for Lower_1_other
unique(MOAB_trimmed_data$Lower_1_other)

# no typos

```

```{r MOAB_clean_Lower_2}

# create a data frame for rows with an "NA" for Lower_2
MOAB_NA_Lower_2 <- subset(MOAB_trimmed_data, is.na(Lower_2))

# there are 38 rows that have an "NA" for Lower_2
# 37 of these rows have "N" for Lower_1
# 1 row has "L-OTHER" for Lower_1 and "EPTO" for Lower_1_other
# we are going to assume all of these should be "N" for Lower_2
MOAB_trimmed_data$Lower_2[is.na(MOAB_trimmed_data$Lower_2)] <- "N"

# look at unique values for Lower_2_other
unique(MOAB_trimmed_data$Lower_2_other)

# no typos

```

```{r MOAB_clean_Lower_3}

# create a data frame for rows with an "NA" for Lower_3
MOAB_NA_Lower_3 <- subset(MOAB_trimmed_data, is.na(Lower_3))

# there are 31 rows that have an "NA" for Lower_3
# all of these rows have an "N" for Lower_2, so we are going to assume they should also have an "N" for Lower_3
MOAB_trimmed_data$Lower_3[is.na(MOAB_trimmed_data$Lower_3)] <- "N"

# look at unique values for Lower_3_other
unique(MOAB_trimmed_data$Lower_3_other)

# no typos

```

```{r MOAB_clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
MOAB_NA_Soil_surface <- subset(MOAB_trimmed_data, is.na(Soil_surface))

# there are 26 rows that have an "NA" for Soil_surface

# after checking Soil_surface values for surrounding points, we will assume these rows should have the default "SOIL" value for Soil_surface
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB104" & MOAB_trimmed_data$Remarks == "3", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB118" & MOAB_trimmed_data$Remarks == "26", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB119" & MOAB_trimmed_data$Remarks == "5", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB128" & MOAB_trimmed_data$Remarks == "9", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB128" & MOAB_trimmed_data$Remarks == "16", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB137" & MOAB_trimmed_data$Remarks == "48", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB148" & MOAB_trimmed_data$Remarks == "7", "Soil_surface"] <- "SOIL"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB148" & MOAB_trimmed_data$Remarks == "26", "Soil_surface"] <- "SOIL"

# the remaining rows with "NA" for Soil_surface are from transect MOAB135 and should be bedrock (expidated data collection due to storm)
MOAB_trimmed_data$Soil_surface[is.na(MOAB_trimmed_data$Soil_surface)] <- "BEDROCK"

# look at unique values for Soil_surface_other
unique(MOAB_trimmed_data$Soil_surface_other)

# fix typos

# "MOSE" should probably be "MOSS"
MOAB_trimmed_data$Soil_surface_other[MOAB_trimmed_data$Soil_surface_other == "MOSE"] <- "MOSS"

# "AHILL" should be "ANTHILL"
MOAB_trimmed_data$Soil_surface_other[MOAB_trimmed_data$Soil_surface_other == "AHILL"] <- "ANTHILL"

```

```{r MOAB_clean_Microsite}

# create a data frame for rows with an "NA" for Microsite
MOAB_NA_Microsite <- subset(MOAB_trimmed_data, is.na(Microsite))

# there are 32 rows that have an "NA" for Microsite
# one of these rows has "L-SHRUB" for Top_layer, so we will assume the Microsite should be "CANOPY"
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "28", "Microsite"] <- "CANOPY"
# all the other rows either have "N" for Top_layer or "L-FORB" or "L-PGRASS", so we will assume a Microsite of "OPEN" for all of them
MOAB_trimmed_data$Microsite[is.na(MOAB_trimmed_data$Microsite)] <- "OPEN"

# look at unique values for Microstie_other
unique(MOAB_trimmed_data$Microsite_other)

# one row has a Microsite_other value of "0" - this should just be "NA"
MOAB_trimmed_data$Microsite_other[MOAB_trimmed_data$Microsite_other == "0"] <- NA

```

```{r MOAB_clean_Top_height}

# create a data frame for rows with an "NA" for Microsite
MOAB_NA_Top_height <- subset(MOAB_trimmed_data, is.na(Top_height))

# there are 29 rows that have an "NA" for Top_height
# one of these rows is missing quite a bit of other data (e.g., litter type/depths), so we will delete it from the data frame (it is from the transect that was cut short by the storm)
MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "28"),]
# the rest all have "N" for Top_layer, so the Top_height values should be "0"
MOAB_trimmed_data$Top_height[is.na(MOAB_trimmed_data$Top_height)] <- 0

```

```{r MOAB_clean_Litter}

# create a data frame for rows with an "NA" for Litter
MOAB_NA_Litter <- subset(MOAB_trimmed_data, is.na(Litter))

# there are 23 rows that have an "NA" for Litter

# they all have litter depth measurements of 0 (or "NA"), so they should have "N" for Litter
MOAB_trimmed_data$Litter[is.na(MOAB_trimmed_data$Litter)] <- "N"

# look at unique values for Litter_other
unique(MOAB_trimmed_data$Litter_other)

# "LT-H_DUNG" should just be "H_DUNG" for consistency
MOAB_trimmed_data$Litter_other[MOAB_trimmed_data$Litter_other == "LT-H_DUNG"] <- "H_DUNG"

# create a data frame for rows with "NA" values for litter depth measurements
MOAB_NA_Litter_depth <- subset(MOAB_trimmed_data,
  is.na(Litter_depth_center) |
  is.na(Litter_depth_north) |
  is.na(Litter_depth_east) |
  is.na(Litter_depth_south) |
  is.na(Litter_depth_west))

# there are 35 rows that have "NA" for at least one litter depth measurement

# make sure litter depth values are numeric
MOAB_trimmed_data$Litter_depth_center <- as.numeric(MOAB_trimmed_data$Litter_depth_center)
MOAB_trimmed_data$Litter_depth_north <- as.numeric(MOAB_trimmed_data$Litter_depth_north)
MOAB_trimmed_data$Litter_depth_east <- as.numeric(MOAB_trimmed_data$Litter_depth_east)
MOAB_trimmed_data$Litter_depth_south <- as.numeric(MOAB_trimmed_data$Litter_depth_south)
MOAB_trimmed_data$Litter_depth_west <- as.numeric(MOAB_trimmed_data$Litter_depth_west)

# 5 of these rows have "LT-H" for Litter, so we will take the average litter depth from the other measurements and paste it into the missing measurement (but one needs to be removed because there are no depth measurements at all)
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB106" & MOAB_trimmed_data$Remarks == "45", "Litter_depth_north"] <- 4
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB116" & MOAB_trimmed_data$Remarks == "7", "Litter_depth_east"] <- 3
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "25", "Litter_depth_center"] <- 0
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "25", "Litter_depth_north"] <- 0
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "25", "Litter_depth_south"] <- 0
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "25", "Litter_depth_west"] <- 0
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB149" & MOAB_trimmed_data$Remarks == "50", "Litter_depth_center"] <- 5
MOAB_trimmed_data[MOAB_trimmed_data$Transect=="MOAB149" & MOAB_trimmed_data$Remarks == "50", "Litter_depth_north"] <- 1

MOAB_trimmed_data <- MOAB_trimmed_data[!(MOAB_trimmed_data$Transect=="MOAB135" & MOAB_trimmed_data$Remarks == "26"),]

# the remaining 30 rows have "N" or "DUNG" or "LT-W" for Litter
MOAB_trimmed_data$Litter_depth_center[is.na(MOAB_trimmed_data$Litter_depth_center)] <- "0"
MOAB_trimmed_data$Litter_depth_north[is.na(MOAB_trimmed_data$Litter_depth_north)] <- "0"
MOAB_trimmed_data$Litter_depth_east[is.na(MOAB_trimmed_data$Litter_depth_east)] <- "0"
MOAB_trimmed_data$Litter_depth_south[is.na(MOAB_trimmed_data$Litter_depth_south)] <- "0"
MOAB_trimmed_data$Litter_depth_west[is.na(MOAB_trimmed_data$Litter_depth_west)] <- "0"

```

```{r MOAB_clean_Nearby_canopy}

# look at unique values for Nearby_canopy_1
unique(MOAB_trimmed_data$Nearby_canopy_1)

# fix typos
MOAB_trimmed_data$Nearby_canopy_1[MOAB_trimmed_data$Nearby_canopy_1 == "L-PGRASS_ L-SUBS_PATCH"] <- "L-PGRASS_L-SUBS_patch"  
MOAB_trimmed_data$Nearby_canopy_1[MOAB_trimmed_data$Nearby_canopy_1 == "L-PGASS_patch"] <- "L-PGRASS_patch"  

# look at unique values for Nearby_canopy_2
unique(MOAB_trimmed_data$Nearby_canopy_2)

# fix typos
MOAB_trimmed_data$Nearby_canopy_2[MOAB_trimmed_data$Nearby_canopy_2 == "L-PGRASS_PATCH"] <- "L-PGRASS_patch"
MOAB_trimmed_data$Nearby_canopy_2[MOAB_trimmed_data$Nearby_canopy_2 == "L-PGRASS_L-SUBS_PATCH"] <- "L-PGRASS_L-SUBS_patch"

# look at unique values for Nearby_canopy_3
unique(MOAB_trimmed_data$Nearby_canopy_3)

# fix typos 
MOAB_trimmed_data$Nearby_canopy_3[MOAB_trimmed_data$Nearby_canopy_3 == "L-FORB_PATCH"] <- "L-FORB_patch"

# look at unique values for Nearby_canopy_4
unique(MOAB_trimmed_data$Nearby_canopy_4)

# look at unique values for Nearby_canopy_5
unique(MOAB_trimmed_data$Nearby_canopy_5)

```

### Level 2 Data

```{r save_MOAB_L2}

# write a csv file of the L2 transect data, save it locally, then upload it to the appropriate Dropbox folder
write.csv(MOAB_trimmed_data, "/Users/AlexiBesser/Desktop/MOAB_LPI_L2_Spring_2025_Data.csv")

```

# ONAQ

## May 2025

**Personnel:** Alexi Besser, Isabel Torres, Dellena Bloom, Jianghai Peng

**Brief Description of Modified LPI Protocol:** Transects ran S to N and were read on the W side. A pin flag was dropped every 1 m along the transect. All vegetation layers (including canopies) intersecting this pin were noted as dead or living and were recorded to species if known and life-form if not. The top layer height (cm), soil surface texture, litter type(s), litter depth (mm), and microsite were also recorded. A SparkFun RTK Express GNSS was used to record the location of each transect point.

50 m transects:

50 m transects with quadrats every 5 m (50 m, 45 m, 40 m, 35 m, 30 m, 25 m, 20 m, 15 m, 10 m, 5 m):

### Levels 0 and 1 Data

```{r ONAQ}

### LEVEL 0 DATA ###

# specify the Dropbox folder containing L0 transect data (individual csv files) and download it locally
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
# FOR NOW: copying and pasting this link in a web browser will automatically download all the csv files
ONAQ_L0_Dropbox_folder <- 

# specify the local folder containing the downloaded L0 transect data
setwd("/Users/AlexiBesser/Downloads/ONAQ_LPI_L0_Spring_2025_csv_files")
ONAQ_L0_data_files <- dir("/Users/AlexiBesser/Downloads/ONAQ_LPI_L0_Spring_2025_csv_files")


### LEVEL 1 DATA ###

# read in L0 csv files and create a new column for transect ID
ONAQ_L1_data <- read_csv(ONAQ_L0_data_files, id = "Transect") %>%
  mutate(Transect = str_remove_all(Transect, ".csv")) %>%
  mutate(Transect = str_remove_all(Transect, "_1")) %>%
  mutate(Transect = str_remove_all(Transect, "_2"))

```

There is now 1 csv file containing data for all 21 transects. This is considered level 1 data.

**L1 Dropbox Filepath:** ASU NSF Dryland Decomposition \> Field&LabCircle \> Field_Data \> ONAQ \> Transect_LPI_Data \> Level_1 \> Spring_2025 \> csv_file

```{r read_ONAQ_L1}

# read in L1 transect data
# HOW TO: open the appropriate Dropbox folder, click the "Share folder" icon in the upper righthand corner, click "Create and copy link" in the window that pops up, then click "Copy link", paste the link here and replace the "0" at the end with "1"
#SRER_L1_data <- read.csv("")

```

### Clean Level 1 Data

Investigate and rectify duplicate rows.

```{r ONAQ_merge_duplicates}

# create a data frame for duplicate points 
ONAQ_duplicate_points <- ONAQ_L1_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup()

ONAQ_trimmed_data <- ONAQ_L1_data

# there are 569 rows of duplicate points
ONAQ_dup_NAcols_removed <- ONAQ_duplicate_points %>%
  filter(`Instrument Ht` == 1.74) %>%
  select(-c(Top_layer:Quadrat_checklist)) %>%
  merge(ONAQ_duplicate_points %>%
    filter(`Instrument Ht` == 0) %>%
    select(-c(Geometry:`Ortho Height`, `Fix ID`:`Satellites in Use`)), 
    by = c("Transect", "Remarks"))

colnames(ONAQ_dup_NAcols_removed)[colnames(ONAQ_dup_NAcols_removed) == "ID.x"] <- "ID"

colnames(ONAQ_dup_NAcols_removed)[colnames(ONAQ_dup_NAcols_removed) == "Time.x"] <- "Time"

colnames(ONAQ_dup_NAcols_removed)[colnames(ONAQ_dup_NAcols_removed) == "Instrument Ht.x"] <- "Instrument Ht"

ONAQ_dup_NAcols_removed <- ONAQ_dup_NAcols_removed %>%
    select(-c("ID.y", "Time.y", "Instrument Ht.y"))

ONAQ_trimmed_data <- ONAQ_trimmed_data %>%
  filter(!(`Instrument Ht` == 0))

ONAQ_trimmed_data <- ONAQ_trimmed_data %>%
  anti_join(ONAQ_dup_NAcols_removed, by = c("Transect", "Remarks")) %>%
  bind_rows(ONAQ_dup_NAcols_removed)

# create a data frame for duplicate points 
ONAQ_duplicate_points_check <- ONAQ_trimmed_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup()

# there are 8 duplicate points remaining
# ONAQ213 (45): 2 rows with "45" for Remarks column ("19" and "10" in ID column). One of these was a typo and should be "35"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ213" & ONAQ_trimmed_data$ID == "19" & ONAQ_trimmed_data$Remarks == "45", "Remarks"] <- 35

# ONAQ231 (39): 2 rows with "39" for Remarks column ("15" and "14" in ID column). One of these was a typo and should be "38"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ231" & ONAQ_trimmed_data$ID == "15" & ONAQ_trimmed_data$Remarks == "39", "Remarks"] <- 38

# ONAQ242 (35): 2 rows with "35" for Remarks column ("17" and "18" in ID column). ID "18" contains the file name for a photograph ("20250527170620.jpg") - this file name will be copied and pasted into the Photo column for the ID "17" row and the ID "18" row will be deleted.
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ242" & ONAQ_trimmed_data$ID == "17", "Photo"] <- "20250527170620.jpg"
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ242" & ONAQ_trimmed_data$Remarks == "35" & ONAQ_trimmed_data$ID == "18"),]

# ONAQ226 (42): 2 rows with "42" for Remarks column ("49" and "51" in ID column). They are identical, so the first row will be deleted.
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ226" & ONAQ_trimmed_data$Remarks == "42" & ONAQ_trimmed_data$ID == "49"),]

```

Look at unique *Remarks*. Remove rows that are not points on transects but rather notes at the end of transects and rows with "NA" in every column.

```{r ONAQ_clean_Remarks}

# extract and review unique Remarks
unique(ONAQ_trimmed_data$Remarks)

```

There are now 1,000 rows in the trimmed data frame! 20 transects x 50 points per transect = 1,000 data points, so this makes sense, but double-check that there are 50 rows for every transect.

```{r ONAQ_check_rows_per_transect}

# create a character string of the transects
ONAQ_Transects <- unique(ONAQ_trimmed_data$Transect)

# create an empty list for the number of rows per transect
ONAQ_rows_per_transect <- list()

# run a for loop to count the number of rows per transect
for (i in ONAQ_Transects) {
  ONAQ_rows_per_transect[[i]] <- sum(ONAQ_trimmed_data$Transect == i)
  cat("Transect:", i, "- Rows:", ONAQ_rows_per_transect[[i]], "\n")
}

```

Investigate "NA" for *Top_layer* and look at unique values for *Top_layer_other* to check for typos.

```{r ONAQ_clean_Top_layer}

# create a data frame for rows with an "NA" for Top_layer
ONAQ_NA_Top_layer <- subset(ONAQ_trimmed_data, is.na(Top_layer))

# there are 84 rows that have an "NA" for Top_layer

# ONAQ226 (44): has "JUOS" for Top_layer_other, so the Top_layer should be "L-OTHER"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect == "ONAQ226" & ONAQ_trimmed_data$Remarks == "44", "Top_layer"] <- "L-OTHER"

# ONAQ243 (6): has "SP" for Top_layer_other, so the Top_layer should be "L-OTHER"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect == "ONAQ243" & ONAQ_trimmed_data$Remarks == "6", "Top_layer"] <- "L-OTHER"

# the following rows have to be removed because they have values for Top_height but we have no idea what the Top_layer is
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ210" & ONAQ_trimmed_data$Remarks == "5"),]
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ210" & ONAQ_trimmed_data$Remarks == "27"),]
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ228" & ONAQ_trimmed_data$Remarks == "26"),]
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ228" & ONAQ_trimmed_data$Remarks == "27"),]
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ204" & ONAQ_trimmed_data$Remarks == "9"),]
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ232" & ONAQ_trimmed_data$Remarks == "14"),]

# for now, delete rows with an "NA" for Top_height because important information is missing for most other columns for these rows as well
ONAQ_trimmed_data <- ONAQ_trimmed_data[!is.na(ONAQ_trimmed_data$Top_height), ]

# the remaining rows have "0" for Top_height, so we will assume Top_layer should be "N"
ONAQ_trimmed_data$Top_layer[is.na(ONAQ_trimmed_data$Top_layer)] <- "N"

# look at unique values for Top_layer_other
unique(ONAQ_trimmed_data$Top_layer_other)

# fix typos
ONAQ_trimmed_data$Top_layer_other[ONAQ_trimmed_data$Top_layer_other == "JUORS"] <- "JUOS"

```

```{r ONAQ_clean_Lower_1}

# create a data frame for rows with an "NA" for Lower_1
ONAQ_NA_Lower_1 <- subset(ONAQ_trimmed_data, is.na(Lower_1))

# there are 85 rows that have an "NA" for Lower_1

# For now, we are going to assume all of these should be "N" for Lower_1
ONAQ_trimmed_data$Lower_1[is.na(ONAQ_trimmed_data$Lower_1)] <- "N"

# look at unique values for Lower_1_other
unique(ONAQ_trimmed_data$Lower_1_other)

# fix typos
ONAQ_trimmed_data$Lower_1_other[ONAQ_trimmed_data$Lower_1_other == "BRTE30"] <- "BRTE"

```

```{r ONAQ_clean_Lower_2}

# create a data frame for rows with an "NA" for Lower_2
ONAQ_NA_Lower_2 <- subset(ONAQ_trimmed_data, is.na(Lower_2))

# there are 101 rows that have an "NA" for Lower_2

# For now, we are going to assume all of these should be "N" for Lower_2
ONAQ_trimmed_data$Lower_2[is.na(ONAQ_trimmed_data$Lower_2)] <- "N"

# look at unique values for Lower_2_other
unique(ONAQ_trimmed_data$Lower_2_other)

# no typos

```

```{r ONAQ_clean_Lower_3}

# create a data frame for rows with an "NA" for Lower_3
ONAQ_NA_Lower_3 <- subset(ONAQ_trimmed_data, is.na(Lower_3))

# there are 101 rows that have an "NA" for Lower_3
# For now, we are going to assume they should also have an "N" for Lower_3
ONAQ_trimmed_data$Lower_3[is.na(ONAQ_trimmed_data$Lower_3)] <- "N"

# look at unique values for Lower_3_other
unique(ONAQ_trimmed_data$Lower_3_other)

# no typos

```

```{r ONAQ_clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
ONAQ_NA_Soil_surface <- subset(ONAQ_trimmed_data, is.na(Soil_surface))

# there are 69 rows that have an "NA" for Soil_surface

# For now, we will assume these rows should have the default "SOIL" value for Soil_surface
ONAQ_trimmed_data$Soil_surface[is.na(ONAQ_trimmed_data$Soil_surface)] <- "SOIL"

# look at unique values for Soil_surface_other
unique(ONAQ_trimmed_data$Soil_surface_other)

# fix typos

# "PGRASS" should be "L-PGRASS"
ONAQ_trimmed_data$Soil_surface_other[ONAQ_trimmed_data$Soil_surface_other == "PGRASS"] <- "L-PGRASS"

# "D_shrub" should be "D-SHRUB"
ONAQ_trimmed_data$Soil_surface_other[ONAQ_trimmed_data$Soil_surface_other == "D_shrub"] <- "D-SHRUB"

# "MOSE" should be "MOSS"
ONAQ_trimmed_data$Soil_surface_other[ONAQ_trimmed_data$Soil_surface_other == "MOSE"] <- "MOSS"

# "A_HILL" should be "ANTHILL"
ONAQ_trimmed_data$Soil_surface_other[ONAQ_trimmed_data$Soil_surface_other == "A_HILL"] <- "ANTHILL"

```

```{r ONAQ_clean_Microsite}

# create a data frame for rows with an "NA" for Microsite
ONAQ_NA_Microsite <- subset(ONAQ_trimmed_data, is.na(Microsite))

# there are 47 rows that have an "NA" for Microsite
# three of these rows have "L-OTHER" for Top_layer and "JUOS" or "ARTR" for Top_layer_other, so the Microsite should probably be "CANOPY"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ226" & ONAQ_trimmed_data$Remarks == "44", "Microsite"] <- "CANOPY"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ226" & ONAQ_trimmed_data$Remarks == "38", "Microsite"] <- "CANOPY"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ228" & ONAQ_trimmed_data$Remarks == "6", "Microsite"] <- "CANOPY"

# For now, we will assume the other rows should have a Microsite of "OPEN"
ONAQ_trimmed_data$Microsite[is.na(ONAQ_trimmed_data$Microsite)] <- "OPEN"

# look at unique values for Microstie_other
unique(ONAQ_trimmed_data$Microsite_other)

```

```{r ONAQ_clean_Top_height}

# create a data frame for rows with an "NA" for Microsite
ONAQ_NA_Top_height <- subset(ONAQ_trimmed_data, is.na(Top_height))

# there are 0 rows that have an "NA" for Top_height

```

```{r ONAQ_clean_Litter}

# create a data frame for rows with an "NA" for Litter
ONAQ_NA_Litter <- subset(ONAQ_trimmed_data, is.na(Litter))

# there are 24 rows that have an "NA" for Litter

# For the sake of time, we are going to remove these rows for now, as many of them have litter depth measurments
ONAQ_trimmed_data <- ONAQ_trimmed_data[!is.na(ONAQ_trimmed_data$Litter),]

# look at unique values for Litter_other
unique(ONAQ_trimmed_data$Litter_other)

# "W_" should just be "W" for consistency
ONAQ_trimmed_data$Litter_other[ONAQ_trimmed_data$Litter_other == "W_"] <- "W"

# create a data frame for rows with "NA" values for litter depth measurements
ONAQ_NA_Litter_depth <- subset(ONAQ_trimmed_data,
  is.na(Litter_depth_center) |
  is.na(Litter_depth_north) |
  is.na(Litter_depth_east) |
  is.na(Litter_depth_south) |
  is.na(Litter_depth_west))

# there are 8 rows that have "NA" for at least one litter depth measurement

# make sure litter depth values are numeric
ONAQ_trimmed_data$Litter_depth_center <- as.numeric(ONAQ_trimmed_data$Litter_depth_center)
ONAQ_trimmed_data$Litter_depth_north <- as.numeric(ONAQ_trimmed_data$Litter_depth_north)
ONAQ_trimmed_data$Litter_depth_east <- as.numeric(ONAQ_trimmed_data$Litter_depth_east)
ONAQ_trimmed_data$Litter_depth_south <- as.numeric(ONAQ_trimmed_data$Litter_depth_south)
ONAQ_trimmed_data$Litter_depth_west <- as.numeric(ONAQ_trimmed_data$Litter_depth_west)

# For now, we will just assume all the "NA" values should be "0"
ONAQ_trimmed_data$Litter_depth_center[is.na(ONAQ_trimmed_data$Litter_depth_center)] <- "0"
ONAQ_trimmed_data$Litter_depth_north[is.na(ONAQ_trimmed_data$Litter_depth_north)] <- "0"
ONAQ_trimmed_data$Litter_depth_east[is.na(ONAQ_trimmed_data$Litter_depth_east)] <- "0"
ONAQ_trimmed_data$Litter_depth_south[is.na(ONAQ_trimmed_data$Litter_depth_south)] <- "0"
ONAQ_trimmed_data$Litter_depth_west[is.na(ONAQ_trimmed_data$Litter_depth_west)] <- "0"

```

```{r ONAQ_clean_Nearby_canopy}

# look at unique values for Nearby_canopy_1
unique(ONAQ_trimmed_data$Nearby_canopy_1)

# fix typos
ONAQ_trimmed_data$Nearby_canopy_1[ONAQ_trimmed_data$Nearby_canopy_1 == "L-PGRASS_PATCH"] <- "L-PGRASS_patch"  

# look at unique values for Nearby_canopy_2
unique(ONAQ_trimmed_data$Nearby_canopy_2)

# fix typos
ONAQ_trimmed_data$Nearby_canopy_2[ONAQ_trimmed_data$Nearby_canopy_2 == "L-BRTE_PGRASS_D-BRTE_DPGRASS_patch"] <- "L-BRTE_L-PGRASS_D-BRTE_D-PGRASS_patch"
ONAQ_trimmed_data$Nearby_canopy_2[ONAQ_trimmed_data$Nearby_canopy_2 == "LPGRASS_DPGRASS_L-FORB_patch"] <- "L-PGRASS_D-PGRASS_L-FORB_patch"

# look at unique values for Nearby_canopy_3
unique(ONAQ_trimmed_data$Nearby_canopy_3)

# fix typos 
ONAQ_trimmed_data$Nearby_canopy_3[ONAQ_trimmed_data$Nearby_canopy_3 == "LPGRASS_patch"] <- "L-PGRASS_patch"

# look at unique values for Nearby_canopy_4
unique(ONAQ_trimmed_data$Nearby_canopy_4)

# look at unique values for Nearby_canopy_5
unique(ONAQ_trimmed_data$Nearby_canopy_5)

```

### Level 2 Data

```{r save_ONAQ_L2}

# write a csv file of the L2 transect data, save it locally, then upload it to the appropriate Dropbox folder
write.csv(ONAQ_trimmed_data, "/Users/AlexiBesser/Desktop/ONAQ_LPI_L2_Spring_2025_Data.csv")

```

# 
