---
title: "ONAQ October-November 2024 Transects"
author: "Alexi C. Besser"
format:
  html:
    embed-resources: true
    code-background: true
    code-line-numbers: true
toc: true
editor: visual
---

# READ ME

Analyses of modified LPI transect data collected at the Onaqui National Ecological Observatory Network Site 4-7 November 2024. 50 m transects:

# Load Packages

```{r load_packages}

library(readr)
library(plyr)
library(dplyr)
library(stringr)
library(ggplot2)

```

# Read In and Clean Up Data

There are 30 csv files each containing data for half a transect.

```{r import_data}

### LEVEL 0 DATA ###
# specify folder containing trasect data (individual csv files are considered Level 0 data)
ONAQ_iPad_data_files <- dir("~/Downloads/ONAQ_Final_Fall_2024/iPad")

# read in csv files and create a new column for transect id
# set working directory
setwd("~/Downloads/ONAQ_Final_Fall_2024/iPad")
ONAQ_iPad_data <- read_csv(ONAQ_iPad_data_files, id = "Transect") %>%
   mutate(Transect = str_remove_all(Transect, "_B.csv"))

# specify folder containing trasect data (individual csv files are considered Level 0 data)
ONAQ_Tablet_data_files <- dir("~/Downloads/ONAQ_Final_Fall_2024/Tablet")

# read in csv files and create a new column for transect id
setwd("~/Downloads/ONAQ_Final_Fall_2024/Tablet")
ONAQ_Tablet_data <- read_csv(ONAQ_Tablet_data_files, id = "Transect") %>%
   mutate(Transect = str_remove_all(Transect, "_A.csv"))

# need to remove "Speed" and "Bearing" columns from the iPad data frame to merge it with the Tablet data frame
ONAQ_Tablet_data <- ONAQ_Tablet_data %>% select(-Speed, -Bearing)
  
# set working directory
setwd("~/Downloads/ONAQ_Final_Fall_2024")

### LEVEL 1 DATA ###
### ONAQ_all_data ###

# merge iPad and Tablet data frames
ONAQ_all_data <- rbind(ONAQ_iPad_data, ONAQ_Tablet_data)

# write a csv file of the compiled (Level 1) data
write.csv(ONAQ_all_data, "~/Desktop/ONAQ_LPI_L1_Fall_2024_Data.csv")

```

```{r remove_NA}

# if all the transect data columns are empty, delete
ONAQ_trimmed_data <- ONAQ_all_data[!(is.na(ONAQ_all_data$Top_layer) & is.na(ONAQ_all_data$Top_layer_other) & is.na(ONAQ_all_data$Lower_1) & is.na(ONAQ_all_data$Lower_1_other) & is.na(ONAQ_all_data$Lower_2) & is.na(ONAQ_all_data$Lower_2_other) & is.na(ONAQ_all_data$Lower_3) & is.na(ONAQ_all_data$Lower_3_other) & is.na(ONAQ_all_data$Litter) & is.na(ONAQ_all_data$Litter_other) & is.na(ONAQ_all_data$Soil_surface) & is.na(ONAQ_all_data$Microsite) & is.na(ONAQ_all_data$Microsite_other)),] 

# create a data frame for duplicate points 
ONAQ_duplicate_points <- ONAQ_trimmed_data %>%
  group_by(Transect, Remarks) %>%
  filter(n() > 1) %>%
  ungroup() 

# delete rows that have duplicates with conflicting data
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ020" & ONAQ_trimmed_data$Remarks == "16"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ026" & ONAQ_trimmed_data$Remarks == "20"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ011" & ONAQ_trimmed_data$Remarks == "28"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ045" & ONAQ_trimmed_data$Remarks == "36"),]

```

Investigate "NA" for Top_layer. And look at unique values for Top_layer_other to check for typos.

```{r clean_Top_layer}

# create a data frame for rows with an "NA" for Top_layer
NA_Top_layer_tab <- subset(ONAQ_trimmed_data, is.na(Top_layer))

# there is one row with an "NA" for Top_layer
# ONAQ003 (Remarks "6") has an "L-OTHER" for Lower_1 and "JUOS" for Lower_1_other
# because ONAQ003 (Remarks "7") has an "L-OTHER" for Top_layer and "JUOS" for Top_layer_other plus a similar Top_height, we will assume this should be the top layer
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ003" & ONAQ_trimmed_data$Remarks == "6", "Top_layer"] <- "L-OTTHER"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ003" & ONAQ_trimmed_data$Remarks == "6", "Top_layer_other"] <- "JUOS"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ003" & ONAQ_trimmed_data$Remarks == "6", "Lower_1"] <- "N"
ONAQ_trimmed_data[ONAQ_trimmed_data$Transect=="ONAQ003" & ONAQ_trimmed_data$Remarks == "6", "Lower_1_other"] <- "N"

# look at unique values for Top_layer_other
Top_layer_other <- unique(ONAQ_trimmed_data$Top_layer_other)
print(Top_layer_other)

# fix typos

# "JKUOS" appears once for transect ONAQ003 - this should be "JUOS"
ONAQ_trimmed_data$Top_layer_other[ONAQ_trimmed_data$Top_layer_other=="JKUOS"] <- "JUOS"

# "JUOAs" appears once for transect ONAQ030 - this should be "JUOS"
ONAQ_trimmed_data$Top_layer_other[ONAQ_trimmed_data$Top_layer_other=="JUOAs"] <- "JUOS"

# "JUKOS" appears once for transect ONAQ030 - this should be "JUOS"
ONAQ_trimmed_data$Top_layer_other[ONAQ_trimmed_data$Top_layer_other=="JUKOS"] <- "JUOS"

# 
ONAQ_trimmed_data$Top_layer_other[ONAQ_trimmed_data$Top_layer_other=="JUKOS"] <- "JUOS"

```

```{r clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
NA_Soil_surface_tab <- subset(ONAQ_trimmed_data, is.na(Soil_surface))

# there are is one row that have an "NA" for Soil_surface

# default Soil_surface value was "SOIL" - let's go ahead and assume any "NA" values in this column should be "SOIL"
# if Soil_surface is "NA" paste "SOIL"
ONAQ_trimmed_data$Soil_surface[is.na(ONAQ_trimmed_data$Soil_surface)] <- "SOIL"

# rename "CRST" to "CRUST"
ONAQ_trimmed_data$Soil_surface[ONAQ_trimmed_data$Soil_surface=="CRST"] <- "CRUST"

# look at unique values for Soil_surface_other
Soil_surface_other <- unique(ONAQ_trimmed_data$Soil_surface_other)
print(Soil_surface_other)

# fix typos
ONAQ_trimmed_data$Soil_surface_other[ONAQ_trimmed_data$Soil_surface_other=="DPGRASS"] <- "D-PGRASS"

```

Investigate "NA" for Litter. And look at unique values for Litter_other to check for typos.

```{r clean_Litter}

# create a data frame for rows with an "NA" for Litter
NA_Litter_tab <- subset(ONAQ_trimmed_data, is.na(Litter))

# there is 1 row that has an "NA" for Litter, it also has zeros for all the litter depths
# if Litter is "NA" paste "N"
ONAQ_trimmed_data$Litter[is.na(ONAQ_trimmed_data$Litter)] <- "N"

# rename "N" to "NONE"
ONAQ_trimmed_data$Litter[ONAQ_trimmed_data$Litter=="N"] <- "NONE"

# create a new data frame for rows that have "NONE" for Litter but still have a litter depth measurement
litter_none_df <- ONAQ_trimmed_data %>%
  filter(Litter == "NONE" & Litter_depth_center > 0 & Litter_depth_north > 0
         & Litter_depth_east > 0 & Litter_depth_south > 0 & Litter_depth_west > 0)
# there are none

ONAQ_trimmed_data$Litter_depth_center[is.na(ONAQ_trimmed_data$Litter_depth_center)] <- 0
ONAQ_trimmed_data$Litter_depth_north[is.na(ONAQ_trimmed_data$Litter_depth_north)] <- 0
ONAQ_trimmed_data$Litter_depth_east[is.na(ONAQ_trimmed_data$Litter_depth_east)] <- 0
ONAQ_trimmed_data$Litter_depth_south[is.na(ONAQ_trimmed_data$Litter_depth_south)] <- 0
ONAQ_trimmed_data$Litter_depth_west[is.na(ONAQ_trimmed_data$Litter_depth_west)] <- 0

# make sure Litter_depth values are numeric
ONAQ_trimmed_data$Litter_depth_center <- as.numeric(ONAQ_trimmed_data$Litter_depth_center)
ONAQ_trimmed_data$Litter_depth_north <- as.numeric(ONAQ_trimmed_data$Litter_depth_north)
ONAQ_trimmed_data$Litter_depth_east <- as.numeric(ONAQ_trimmed_data$Litter_depth_east)
ONAQ_trimmed_data$Litter_depth_south <- as.numeric(ONAQ_trimmed_data$Litter_depth_south)
ONAQ_trimmed_data$Litter_depth_west <- as.numeric(ONAQ_trimmed_data$Litter_depth_west)

```

Investigate "NA" for Microsite.

```{r clean_Microsite}

# create a data frame for rows with an "NA" for Microsite
NA_Microsite_tab <- subset(ONAQ_trimmed_data, is.na(Microsite))

# there are 5 rows that have an "NA" for Microsite
# Microsite_other also has an "NA" for all of these rows
# 4 of these rows have no top layer
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ027" & ONAQ_trimmed_data$Remarks == "13"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ027" & ONAQ_trimmed_data$Remarks == "16"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ042" & ONAQ_trimmed_data$Remarks == "4"),]

ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ045" & ONAQ_trimmed_data$Remarks == "20"),]

# has a Top_layer of "L-OTHER" but an "NA" in the Top_layer_other column... need to delete
ONAQ_trimmed_data <- ONAQ_trimmed_data[!(ONAQ_trimmed_data$Transect=="ONAQ049" & ONAQ_trimmed_data$Remarks == "9"),]

```

Investigate Top_height.

```{r clean_Top_height}

# if Top_height is "NA" paste "0"
ONAQ_trimmed_data$Top_height[is.na(ONAQ_trimmed_data$Top_height)] <- 0

# make sure Top_height values are numeric
ONAQ_trimmed_data$Top_height <- as.numeric(ONAQ_trimmed_data$Top_height)

```

Investigate Soil_surface.

```{r clean_Soil_surface}

# create a data frame for rows with an "NA" for Soil_surface
NA_Soil_tab <- subset(ONAQ_trimmed_data, is.na(Soil_surface))

# there is one row that has an "NA" for Soil_surface

# default Soil_surface value was "SOIL" - let's go ahead and assume any "NA" values in this column should be "SOIL"
# if Soil_surface is "NA" paste "SOIL"
ONAQ_trimmed_data$Soil_surface[is.na(ONAQ_trimmed_data$Soil_surface)] <- "SOIL"

```

# Top Layer Plots

```{r Top_layer_transects}

# first get counts, proportions, and percentages of each top layer type per transect
TopLayer <- ONAQ_trimmed_data %>%
  group_by(Top_layer, Transect) %>%
  dplyr::summarize(count = n(), .groups = "drop") %>%
  group_by(Transect) %>%
  dplyr::mutate(proportion = count / sum(count)) %>%
  dplyr::mutate(percent = proportion*100)

# make a bar plot
TopLayerPlot <- ggplot(TopLayer, aes(x = Transect, y = percent, fill = Top_layer)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Transect", y = "Frequency (%)", fill = "Top Layer") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("navy", "#E69F00", "#56B4E9", "#009E73", "#F0E442",
                               "#0072B2", "#D55E00", "darkgreen", "#CC79A7", "coral4",
                               "#b66dff", "#999999", "seashell", "red", "#000000")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

TopLayerPlot

```

```{r Top_layer_other_transects}

# first, get counts, proportions, and percentages of each Top_layer_other category per transect
TopLayerOther <- ONAQ_trimmed_data %>%
  group_by(Top_layer_other) %>%
  dplyr::summarize(count = n(), .groups = "drop") %>%
  group_by(Transect) %>%
  dplyr::mutate(proportion = count / sum(count)) %>%
  dplyr::mutate(percent = proportion*100)

# make a bar plot
TopLayerOtherPlot <- ggplot(TopLayerOther, aes(x = Transect, y = percent, fill = Top_layer_other)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Transect", y = "Frequency (%)", fill = "Top Layer Other") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("navy", "#E69F00", "#56B4E9", "#009E73", "#F0E442",
                               "#0072B2", "#D55E00", "darkgreen", "#CC79A7", "coral4",
                               "#b66dff", "#999999", "seashell", "red", "darkmagenta",
                               "olivedrab3", "#000000"), 
                    name = NULL, 
                    breaks = c("BOGR", "GUSA", "ARTR", "JUOS", "EPTO", "ACHY",
                               "PI", "SAVE", "KRLA", "SATR", "EPVI", "NONE"),
                    labels = c("BOGR", "GUSA", "ARTR", "JUOS", "EPTO", "ACHY",
                               "PI", "SAVE", "KRLA", "SATR", "EPVI", "NONE")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

TopLayerOtherPlot

```

# Soil Surface Plots

## Soil Surface Frequencies Across Transects

### Soil_surface Plot

```{r Soil_surface_transects}

# first, get counts, proportions, and percentages of each soil surface type per transect
SoilSurface <- ONAQ_trimmed_data %>%
  group_by(Soil_surface, Transect) %>%
  dplyr::summarize(count = n(), .groups = "drop") %>%
  group_by(Transect) %>%
  dplyr::mutate(proportion = count / sum(count)) %>%
  dplyr::mutate(percent = proportion*100)

# make a bar plot
SoilSurfacePlot <- ggplot(SoilSurface, aes(x = Transect, y = percent, fill = Soil_surface)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Transect", y = "Frequency (%)", fill = "Soil Surface") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("gray20", "#E69F00", "#56B4E9", "coral4", "#D55E00",
                               "burlywood", "#999999", "burlywood4",
                               "darkslateblue")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

SoilSurfacePlot

```

## Top Layer Frequencies Across Soil Surfaces

```{r Top_layer_Soil_surface}

# first, get counts, proportions, and percentages of each top layer type per transect
SoilTopLayer <- ONAQ_trimmed_data %>%
  group_by(Top_layer_other, Soil_surface) %>%
  dplyr::summarize(count = n(), .groups = "drop") %>%
  group_by(Soil_surface) %>%
  dplyr::mutate(proportion = count / sum(count)) %>%
  dplyr::mutate(percent = proportion*100)

# make a bar plot
SoilTopLayerPlot <- ggplot(SoilTopLayer, aes(x = Soil_surface, y = percent, fill = Top_layer_other)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Soil Surface", y = "Frequency (%)", fill = "Top Layer") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("burlywood4", "#999999", "burlywood", "mediumpurple4",
                               "coral4", "#D55E00", "#E69F00", "#F0E442", "#009E73",
                               "#CC79A7", "#0072B2", "cornsilk2", "darkgreen",
                               "olivedrab3", "slategray1", "violetred1", "#000000")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

SoilTopLayerPlot

```

# Litter Plots

## Litter Types and Depths Across Transects

### Litter Plot

```{r Litter_transects}

# first, get counts, proportions, and percentages of each litter type per transect
LitterTransect <- ONAQ_trimmed_data %>%
  group_by(Litter, Transect) %>%
  dplyr::summarize(totdepth_center = sum(Litter_depth_center),
                   totdepth_north = sum(Litter_depth_north),
                   totdepth_east = sum(Litter_depth_east),
                   totdepth_south = sum(Litter_depth_south),
                   totdepth_west = sum(Litter_depth_west),
                   n = n(), .groups = "drop") %>%
  group_by(Transect) %>%
  dplyr::mutate(proportion = n / sum(n)) %>%
  dplyr::mutate(percent = proportion*100)

# make a bar plot
LitterPlot <- ggplot(LitterTransect, aes(x = Transect, y = percent, fill = Litter)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Transect", y = "Frequency (%)", fill = "Litter") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("coral4", "#009E73", "burlywood4",
                               "black", "orange", "#999999")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

LitterPlot

```

### Litter_depth Plot

Make plot that shows litter type and cumulative depth across transects.

```{r Litter_depth_transects}

# make a bar plot
LitterDepthPlot <- ggplot(LitterTransect, aes(x = Transect, y = totdepth_center,
                                           fill = Litter)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Transect", y = "Cumulative Litter Depth (mm)", fill = "Litter Type") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("coral4", "burlywood", "#009E73", "burlywood4",
                               "black", "#999999"), 
                    name = NULL, 
                    breaks = c("DUNG", "LT-C", "LT-H", "LT-W", "NONE",
                               "OTHER"),
                    labels = c("Dung", "Cactus Litter", "Nonwoody Litter",
                               "Woody Litter", "No Litter", "Other Litter")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = c(.2,.7),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

LitterDepthPlot

```

## Litter Types and Depths Across Soil Surfaces

Make a plot of litter type and cumulative depth as a function of soil surface.

```{r Litter_depth_Soil_surface}

# first, get total depths of each litter type per soil surface type
LitterSoil <- ONAQ_trimmed_data %>%
  group_by(Soil_surface, Litter) %>%
  dplyr::summarize(totdepth_center = sum(Litter_depth_center),
                   totdepth_north = sum(Litter_depth_north),
                   totdepth_east = sum(Litter_depth_east),
                   totdepth_south = sum(Litter_depth_south),
                   totdepth_west = sum(Litter_depth_west),
                   n = n(), .groups = "drop")

# make a bar plot
LitterSoilPlot <- ggplot(LitterSoil, aes(x = Soil_surface, y = totdepth_center,
                                         fill = Litter)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Soil Surface", y = "Cumulative Litter Depth (mm)", fill = "Litter Type") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  scale_fill_manual(values = c("coral4", "burlywood", "#009E73", "burlywood4",
                               "black", "#999999"), 
                    name = NULL, 
                    breaks = c("DUNG", "LT-C", "LT-H", "LT-W", "NONE",
                               "OTHER"),
                    labels = c("Dung", "Cactus Litter", "Nonwoody Litter",
                               "Woody Litter", "No Litter", "Other Litter")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

LitterSoilPlot

```

## Litter Types and Depths Across Microsites

```{r Litter_depth_Microsite}

# look at unique values for Litter_other
Litter_other <- unique(ONAQ_trimmed_data$Litter_other)
print(Litter_other)

# Make a new column
ONAQ_trimmed_data$Litter_Plotting <- ONAQ_trimmed_data$Litter

# if Litter_other is "H_W" paste "Nonwoody and Woody Litter"
for (i in 1:nrow(ONAQ_trimmed_data)) {
  if (!is.na(ONAQ_trimmed_data$Litter_other[i]) &&
      ONAQ_trimmed_data$Litter_other[i] == "H_W")
    {ONAQ_trimmed_data$Litter_Plotting[i] <- "Nonwoody and Woody Litter"}
}

# if Litter_other is "H_W_B" paste "Nonwoody and Woody Litter"
for (i in 1:nrow(ONAQ_trimmed_data)) {
  if (!is.na(ONAQ_trimmed_data$Litter_other[i]) &&
      ONAQ_trimmed_data$Litter_other[i] == "H_W_B")
    {ONAQ_trimmed_data$Litter_Plotting[i] <- "Nonwoody and Woody Litter"}
}

# if Litter_other is "H_B" paste "Nonwoody and Woody Litter"
for (i in 1:nrow(ONAQ_trimmed_data)) {
  if (!is.na(ONAQ_trimmed_data$Litter_other[i]) &&
      ONAQ_trimmed_data$Litter_other[i] == "H_W")
    {ONAQ_trimmed_data$Litter_Plotting[i] <- "Nonwoody and Woody Litter"}
}

# first, get counts, proportions, and percentages of each litter type per top layer category
LitterMic <- ONAQ_trimmed_data %>%
  group_by(Litter_Plotting, Microsite) %>%
  dplyr::summarize(totdepth_center = sum(Litter_depth_center),
                   totdepth_north = sum(Litter_depth_north),
                   totdepth_east = sum(Litter_depth_east),
                   totdepth_south = sum(Litter_depth_south),
                   totdepth_west = sum(Litter_depth_west),
                   n = n(),
                   mean_totdepth_all = (sum(Litter_depth_center) +
                     sum(Litter_depth_north) + sum(Litter_depth_east) +
                     sum(Litter_depth_south) + sum(Litter_depth_west))/(5*n),
                   .groups = "drop")

# next, calculate the total number of observations per microsite
MicroFreq <- ONAQ_trimmed_data %>%
  group_by(Microsite) %>%
  dplyr::summarize(nMic = n(), .groups = "drop")

# merge the MicroFreq data frame with the LitterStatsMicro data frame
LitterMic <- LitterMic %>%
  left_join(MicroFreq, by = c("Microsite"))

# calculate the total number of observations
total_mic_counts <- sum(LitterMic$n)

# calculate the maximum sum of mean litter depths for a top layer category
sum_mean_depth_max_mic <- LitterMic %>%
  group_by(Microsite) %>%
  summarize(sum_mean_depth = sum(mean_totdepth_all)) %>%
  summarize(sum_mean_depth_max = max(sum_mean_depth)) %>%
  pull(sum_mean_depth_max)

# calculate the maximum total number of observations per microsite
mic_freq_max <- max(LitterMic$nMic)

LitterMicPlot <- ggplot(LitterMic, aes(x = Microsite, y = mean_totdepth_all,
                                       fill = Litter_Plotting)) +
  guides(fill = guide_legend(title.position = "top")) +
  labs(x = "Microsite", y = "Mean Litter Depth (mm)", fill = "Litter Type") +
  geom_bar(stat = "identity", position = "stack", color = "black", linewidth = 0.25) +
  geom_point(aes(y = nMic*(sum_mean_depth_max_mic/mic_freq_max), fill = "n"),
             size = 3, pch = 21, color = "black") +
  scale_fill_manual(values = c("burlywood", "#009E73", "#006B4E", "burlywood4",
                               "saddlebrown", "#999999", "black", "navy"), 
                    name = NULL, 
                    breaks = c("LT-C", "LT-H", "Nonwoody and Woody Litter",
                               "LT-W", "DUNG", "OTHER", "NONE", "n"),
                    labels = c("Cactus Litter", "Nonwoody Litter",
                               "Nonwoody and Woody Litter", "Woody Litter",
                               "Dung", "Other Litter", "No Litter", "Count")) +
  scale_y_continuous(
    name = "Mean Litter Depth (mm)",
    limits = c(0, 40),
    sec.axis = sec_axis(~ . * (mic_freq_max/sum_mean_depth_max_mic), 
                        name = "Number of Observations")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.position = (c(0.6, 0.7)),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75)) +
  ggtitle(label = "ONAQ", subtitle = "October-November 2024") +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, hjust = 0.5))

LitterMicPlot

ggsave("~/Desktop/ONAQ_Microsite_Mean_Litter.pdf", LitterMicPlot, width = 10,
       height = 6)

# A LOT of the "OTHER" litter type is actually just a combination of litter types
# need to work on some code to categorize the "OTHER" litter types better

```

```{r}

LitterTransect_all <- ONAQ_trimmed_data %>%
  group_by(Transect) %>%
  dplyr::summarize(totdepth = sum(Litter_depth_center), n = n(), 
                   elev = mean(Elevation), .groups = "drop")

elev_lm <- lm(totdepth ~ elev, data = LitterTransect_all)
summary(elev_lm)

elev_depth <- ggplot(LitterTransect_all, aes(x = elev, y = totdepth)) +
  geom_point(pch = 19, size = 5) +
  stat_smooth(method = "lm", formula = y ~ x, fullrange = TRUE, color = "black") +
  stat_poly_eq(use_label("eq"), formula = y ~ x) +
  stat_poly_eq(use_label("R2", "P"), formula = y ~ x, label.y = 0.9) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(x = "Elevation (m)", y = "Litter Depth (mm)", fill = "Litter Type") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(
    axis.text.x = element_text(color="black", size=14),
    axis.text.y = element_text(color="black", size=14),
    legend.title = element_text(color="black", size=14),
    legend.text = element_text(color="black", size=14),
    legend.position = "right",
    axis.title.x = element_text(color="black", size=16),
    axis.title.y = element_text(color="black", size=16, angle=90),
    panel.grid.major = element_blank(), # switch off major grid lines
    panel.grid.minor = element_blank(), # switch off minor grid lines
    panel.border = element_rect(colour = "black", fill=NA, linewidth = 1),
    axis.ticks = element_line(color = "black", linewidth = 0.75))

elev_depth

```
